---
title: "Reproducible Research: Peer Assessment 1"
============================================================
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r,echo=TRUE}

activity <- read.csv("activity.csv")

```



## What is mean total number of steps taken per day?

```{r fig.width=7, fig.height=6 ,echo=TRUE}
step_by_day <- by(activity$steps, activity$date, sum, na.rm=T)
barplot(step_by_day, ylab="steps", xlab="days", main="Total number of steps taken per day")
```
**median** and **mean** total number of steps taken per day

```{r,echo=TRUE}

summary(activity$steps, activity$date, na.rm=T)[3:4]

```

## What is the average daily activity pattern?
```{r fig.width=7, fig.height=6 ,echo=TRUE}

activity$interval <- sprintf("%04d", activity$interval)
int_mean <- tapply(activity$steps, activity$interval, mean, na.rm=T)
names(int_mean) <- unique(activity$interval)
plot(unique(activity$interval),int_mean, type="l", ylab="steps", xlab="time interval", main="Average number of steps taken per time interval")
```

interval with **max** numm of steps
```{r,echo=TRUE}

int_mean[which(int_mean %in% max(int_mean))]
``` 


## Imputing missing values

total number of **missing values** in the dataset
```{r,echo=TRUE}

sum(is.na(activity$steps))

``` 

histogram of the total number of steps taken each day where **NA** have been filled with the mean for 
5-minute interval

```{r fig.width=7, fig.height=6 ,echo=TRUE}
Na_index <- which(is.na(activity$steps))
activity_new <- activity
activity_new$steps[Na_index] <- int_mean[activity$interval[Na_index]]

stepNew_by_day <- by(activity_new$steps, activity_new$date, sum, na.rm=T)
barplot(stepNew_by_day, ylab="steps", xlab="days", main="Total number of steps taken per day")

``` 

**median** and **mean** total number of steps taken per day for the new daraset

```{r,echo=TRUE}

summary(activity_new$steps, activity_new$date, na.rm=T)[3:4]
```

there is no difference for the totla number of steps taken per day between **median** and **mean** of the new datasates (i.e. without NA value) and the original dataset (with NA value)
```{r,echo=TRUE}
summary(activity_new$steps, activity_new$date, na.rm=T)[3:4] - summary(activity$steps, activity$date, na.rm=T)[3:4]

```
It looks like there is not a relevant impact of imputing missing data on the estimates of the total daily number of steps
## Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}

activity_new$date <- as.Date.factor(activity_new$date)
activity_new$weekend <- weekdays(activity_new$date, abbreviate=T)
activity_new$weekend[activity_new$weekend %in% c("lun","mar","mer","gio","ven")] <- "weekday"
activity_new$weekend[activity_new$weekend %in% c("sab","dom")] <- "weekend"
activity_new$weekend <- as.factor(activity_new$weekend)
table(activity_new$weekend)

```

```{r fig.width=7, fig.height=6 ,echo=TRUE}

par(mfrow = c(2, 1))
for (wd in c("weekday", "weekend")) {
int_mean_2 <- tapply(activity_new$steps[activity_new$weekend==wd], activity_new$interval[activity_new$weekend==wd], mean, na.rm=T)
names(int_mean_2) <- unique(activity_new$interval)
plot(unique(activity_new$interval[activity_new$weekend==wd]),int_mean_2, type="l", ylab="steps", xlab="time interval", main=paste("Average number of steps taken in ", wd, " time interval"))
}

``` 

